### 2020-03-01 doc/icfp20/counterpoint.tex
more tweaks

We next discuss a more complex variant of counterpoint, called the
second species.
In second species, we set \emph{two} half notes against every cantus
firmus note.
Figure~\ref{fig:ss} is an example of such two-against-one counterpoint,
again composed for the Frog's song.

@@

In second species, the first and last bars may have a different structure
from middle bars.
More specifically, it is encouraged to begin the counterpoint line with
a half rest and end with a whole note.
Therefore, in our implementation, we reuse the \texttt{PitchInterval} 
type for the first and last bars, and define a new type
\texttt{PitchInterval2} for middle bars:

@@

In second species counterpoint, some parts of the music are applied
the same rules as first species, while other parts are constrained by
new rules.

@@

The beginning of the music is \emph{not} allowed to be the unison,
as it prevents the listener from recognizing the beginning of the
counterpoint line.
We restrict the first interval by defining a new error
\texttt{BeginningError2} and a function \texttt{checkBeginning2}
that works analogously to \texttt{checkBeginning} for first species.

@@

The rhythmic structure of secound species counterpoint gives rise to
the distinction between \emph{strong beats} (the first interval in a bar)
and \emph{weak beats} (the second interval).
Strong beats in middle bars are constrained by the same rules as in
first species.
That is, they must all be consonant, non-unison intervals, and in the 
case of perfect intervals, the motion from the preceding interval (on
the weak beat) must be either contrary or oblique.
We encode these restrictions as the \texttt{checkStrongBeats}
function, which, as the \texttt{checkIntervals} function for first
species, returns a list of found errors.

@@

Compared to strong beasts, weak beats are constrained in a looser
manner.
First, they are allowed to be dissonant if they are created by a
\emph{passing tone}, i.e., a note in the middle of two step-wise
motions in the same direction (as in bars 4-5 of Figure~\ref{fig:ss}).
Second, they are allowed to be the unison if they are left by step in
the opposite direction from their approach (as in bars 5-6 of
Figure~\ref{fig:ss}).
We implement these rules as the \texttt{checkWeakBeats} function,
which checks the validity of every three successive beats in the
middle bars.

@@

Parallel and similar motion towards a perfect interval is prohibited
across bars.
We enforce this rule using the \texttt{checkMotion2} function,
which uses \texttt{expandPitchIntervals2} to convert a list of
\texttt{PitchInterval2} into a list of \texttt{PitchInterval}, and
checks every pair of weak beat and strong beat intervals.

@@

Now we define \texttt{SecondSpecies}, a record type inhabited by
correct second species counterpoint.
As in \texttt{FirstSpecies}, we have three fields holding the musical
content, followed by five fields carrying the proofs of various
properties.

>>>>

We next discuss a more complex variant of counterpoint, called the
second species.
In second species, we set \emph{two} half notes against every cantus
firmus note.
This gives rise to the distinction between \emph{strong} beats (the
first interval in a bar) and \emph{weak beats} (the second interval).
Figure~\ref{fig:ss} is an example of two-against-one counterpoint,
again composed for the Frog's song.

@@

In second species, the first and last bars may have a different structure
from middle bars.
More specifically, it is encouraged to begin the counterpoint line with
a half rest and end with a whole note.
Therefore, in our implementation, we reuse the \texttt{PitchInterval} 
type for the first and last bars, and define a new type
\texttt{PitchInterval2} for middle bars.

@@

The rules for second species counterpoint can be obtained by
tweaking those for first species and adding a few new ones.
Here we go through the rules without showing the corresponding
Agda functions, as they are largely similar to what we have seen
before.

@@

The beginning of the music is \emph{not} allowed to be the unison,
as it prevents the listener from recognizing the beginning of the
counterpoint line.
% We restrict the first interval by defining a new error
% \texttt{BeginningError2} and a function \texttt{checkBeginning2}
% that works analogously to \texttt{checkBeginning} for first species.

@@

Strong beats in middle bars are constrained by the same rules as in
first species: they must all be consonant, non-unison intervals.
% We encode this restriction as the \texttt{checkStrongBeats}
% function, which, as the \texttt{checkIntervals} function for first
% species, returns a list of found errors.

@@

Weak beats are constrained in a looser manner.
First, they are allowed to be dissonant if they are created by a
\emph{passing tone}, i.e., a note in the middle of two step-wise
motions in the same direction (as in bars 4-5 of Figure~\ref{fig:ss}).
Second, they are allowed to be the unison if they are left by step in
the opposite direction from their approach (as in bars 5-6 of
Figure~\ref{fig:ss}).
% We implement these rules as the \texttt{checkWeakBeats} function,
% which checks the validity of every three successive beats in the
% middle bars.

@@

Parallel and similar motion towards a perfect interval is prohibited
across bars.
% We enforce this rule using the \texttt{checkMotion2} function,
% which uses \texttt{expandPitchIntervals2} to convert a list of
% \texttt{PitchInterval2} into a list of \texttt{PitchInterval}, and
% checks every pair of weak beat and strong beat intervals.

@@

Now we define \texttt{SecondSpecies}, a record type inhabited by
correct second species counterpoint.
As in \texttt{FirstSpecies}, we have three fields holding the musical
content, followed by five fields carrying the proofs of the properties
disucssed above.
